Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-07-09 summary.tex
updated document

In the third example, we consider the Karantaka villages data studied by
\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
networks from villages 12, 31, and 46 at the household level. The label
of interest is the religious affiliation. The networks were truncated to
religions ``1'' and ``2'', and vertices of degree 0 were removed.

@@

\caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left). The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:unnamed-chunk-7}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-8}Community detection error rates for identifying household religion.}
\centering
\begin{tabular}[t]{l|r|r|r}
\hline
Network & MM & SSC-ASE & OSC\\
\hline
Village 12 & 0.270 & 0.291 & 0.227\\
\hline
Village 31 & 0.125 & 0.066 & 0.110\\
\hline
Village 46 & 0.052 & 0.463 & 0.078\\
\hline
\end{tabular}
\end{table}

@@

This paper shows the connection between the PABM and the GRDPG, namely
that a PABM graph can be represented as a union of orthogonal subspaces
in an embedding under the GRDPG framework. We then exploited this
relationship to develop community detection and parameter estimation
methods. In fact, we can represent any graph with Bernoulli edges as a
GRDPG, and in the PABM case, it turns out that this relationship leads
to a very straightforward applications of previous work from
\citet{rubindelanchy2017statistical},
\citet{soltanolkotabi2012}, and \citet{jmlr-v28-wang13}, which
lead to asymptotically correct solutions with high probability. Similar
methods can be applied for other models, such as the Nested Block Model
\cite{noroozi2021hierarchy}.

>>>>

In the third example, we consider the Karantaka villages data studied by
\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
networks from villages 12, 31, and 46 at the household level. The label
of interest is the religious affiliation. The networks were truncated to
religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig \ref{fig:households-figure} and Table \ref{tab:households-table}.

@@

\caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left). The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:households-figure}
\end{figure}

@@

\caption{\label{tab:households-table}Community detection error rates for identifying household religion.}
\centering
\begin{tabular}[t]{l|r|r|r}
\hline
Network & MM & SSC-ASE & OSC\\
\hline
Village 12 & 0.270 & 0.291 & 0.227\\
\hline
Village 31 & 0.125 & 0.066 & 0.110\\
\hline
Village 46 & 0.052 & 0.463 & 0.078\\
\hline
\end{tabular}
\end{table}

@@

This paper connects the Popularity Adjusted Block Model to the 
Generalized Random Dot Product Graph and demonstrates how the connection 
can be exploited for community detection and parameter estimation. 
Similar to how, under the GRDPG framework, 
the Stochastic Block Model can be generated by point masses and 
the Degree-Corrected Block Model can be generated by line segments, 
we show that the latent configuration that induces the PABM 
consists of orthogonal subspaces 
with each subspace corresponding to a community. 
This latent structure can be recovered via Adjacency Spectral Embedding. 
We then apply clustering algorithms on the embedding 
that take advantage of this structure: 
Orthogonal Spectral Clustering, which is a novel algorithm, 
and Sparse Subspace Clustering, which has already been applied 
in some way to the PABM but 
without asymptotic guarantees \cite{noroozi2019estimation}. 
Finally, we show that both algorithms result in perfect community detection 
for large graphs. 
These theoretical results are further demonstrated via simulation studies.
